---
title: "Homework 1: Bowling Simulation Part 1"
author: "Amber Potter"
date: "9/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial Comments

For this homework, I chose to use R with the tidyverse and kableExtra packages. I generated the random variables for pins knocked down from uniform distributions from 0 to the number of pins standing before the roll. In this scenario, I am making the assumption that on any given roll, the bowler has a $\frac{1}{number~of~pins~standing ~+ 1}$ chance of knocking down any possible number of pins (including the event that 0 pins are knocked down). Although this leads to increased variability of each frame's score that may not be accurate to the consistency expected from professional bowlers, I found that it might be more reflective of what an inexperienced bowler, like myself, might achieve. A better way to generate the random variables representing the number of pins knocked down might be to assume a bowler knocks down pins in a way that reflects a normal distribution with some mean and variance. For example, the consistency of better bowlers might then be reflected with a mean closer to 10 and a small variance.

My code does not require any parameters. Ideally, because of the reproducible  and viewer friendly formatting I have chosen for my output output, my work would be best viewed by opening this R Markdown file in RStudio and knitting to html. I recommend right clicking on the file link, clicking save link as, then opening the saved Rmd in R, then clicking "Knit" which will automatically create an html file that includes all my code and my final outputs. Although my code can also be executed by either selecting the option to "Run All" or by manually clicking the green run arrows on each code chunk in order, both of these options will produce somewhat faint tables that may be difficult to read depending on a user's R theme. My output has been optimized for html format. For convenience, I have also included a set.seed() function (currently commented out), in case it would be beneficial to be able to repeatedly run my code with the same random seed. Since this function is currently commented out, each run of my code will generate different random variables and thus will result in different final outputs.

The output of my code includes the main table summarizing each frame (represented as a row), which lists the pins knocked down on each roll, the interpretation of the frame, and the number of points earned in that frame. For each frame, there are columns for Roll 1, Roll 2, and Roll 3. The values for these columns only contain numbers if that given roll took place. For example, Frame 10 is the only frame that will have an integer listed in Roll 3. Similarly, if a Strike is earned prior to the Frame 10, Roll 2 of that frame will not display a number. Instead, to denote no roll, "NA" will be displayed. My output also includes an additional value summarizing the Frame Points column of the main table. In the html document, I have formatted this value to appear beneath the Frame Points Column which makes it easily interpreted as the sum of the points earned in each frame listed above it.


# Bowling Simulation

### Packages

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
```

### Set Seed Option

```{r}
#set.seed(21)
```


### Pins Knocked Down

```{r}
frame <- as_tibble(seq(1,10)) %>%
  rename(frame = value)

first_roll <- frame %>%
  mutate(roll_1_pins_down = sample(0:10, 10, replace=TRUE))

roll_1_pins_down_vector <- first_roll$roll_1_pins_down

roll_2_pins_down_vector <- c()


for(i in 1:10){
  roll_2_pins_down_vector <- append(roll_2_pins_down_vector, 
                             sample(0:(10 - first_roll$roll_1_pins_down[i]), 1))
}


  
second_roll <- first_roll %>%
 mutate(roll_2_pins_down = case_when(roll_1_pins_down != 10 
                                     ~ roll_2_pins_down_vector[frame],
                                     roll_1_pins_down == 10 ~ NA_integer_))

third_roll <- second_roll %>%
  mutate(roll_2_pins_down = case_when(frame == 10 & roll_1_pins_down == 10 
                                           ~ sample(0:10, 1),
                                      TRUE ~ roll_2_pins_down),
         roll_3_pins_down = case_when(frame == 10 & roll_1_pins_down == 10 & roll_2_pins_down == 10 
                                           ~ sample(0:10, 1),
                                      frame == 10 & roll_1_pins_down == 10 & roll_2_pins_down != 10 ~ sample(0:(10 - roll_2_pins_down[10]), 1),
                                      frame == 10 & roll_1_pins_down + roll_2_pins_down == 10 ~ sample(0:10, 1),
                                      TRUE ~ NA_integer_))

roll_1_pins_down_vector <- third_roll$roll_1_pins_down
roll_2_pins_down_vector <- third_roll$roll_2_pins_down
roll_3_pins_down_vector <- third_roll$roll_3_pins_down
```


### Frame Interpretation

```{r}
frame_interpretations <- third_roll %>%
  mutate(frame_interpretation = case_when(roll_1_pins_down == 10 ~ "STRIKE",
                                          roll_1_pins_down + roll_2_pins_down == 10 ~ "SPARE",
                                          TRUE ~ "OPEN"))
```


### Frame Points

```{r}
points <- frame_interpretations %>%
  mutate(frame_points = case_when(frame_interpretation == "OPEN" ~ as.numeric(roll_1_pins_down_vector[frame] + roll_2_pins_down_vector[frame]), # open
                      
                                  frame != 10 & frame_interpretation == "SPARE" ~ 10 + lead(roll_1_pins_down, order_by = frame, n = 1), # spare before 10th frame
                                  
                                  frame == 10 & frame_interpretation == "SPARE" ~ 10 + roll_3_pins_down_vector[frame], # spare on 10th frame
                                  
                                  frame < 9 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) == "STRIKE" ~ 20 + lead(roll_1_pins_down, order_by = frame, n = 2), # 2 strikes in a row before 9th
                                  
                                  frame != 10 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) != "STRIKE" ~ 10 + lead(roll_1_pins_down, order_by = frame, n = 1) + lead(roll_2_pins_down, order_by = frame, n = 1), # strike before 10th frame
                                  
                                  frame == 9 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) == "STRIKE" ~ 20 + lead(roll_2_pins_down, order_by = frame, n = 1), # 2 strikes in a row 9th
                                  
                                  frame = 10 & frame_interpretation == "STRIKE" ~ 10 + roll_2_pins_down_vector[10] + roll_3_pins_down_vector[10], # 10th frame strike
                                  ))
```

### Frame Summaries and Total Score

```{r}
frame_summaries <- points %>%
  kable(col.names = c("Frame", "Roll 1 Pins Down", "Roll 2 Pins Down", "Roll 3 Pins Down", "Frame Interpretation", "Frame Points"), align = "r") %>%
  kableExtra::kable_styling("striped")


total_score <- points %>%
  summarize(total_points = sum(frame_points)) %>%
  kable(col.names = "Total Points", align = "r") %>%
  kableExtra::kable_styling("striped")
```

```{r eval = FALSE}
frame_summaries
total_score
```

```{r echo = FALSE}
frame_summaries
```


```{r echo = FALSE}
total_score
```



