---
title: "lecture-notes-0908"
author: "Amber Potter"
date: "9/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Bowling Simulation Part 2

1. Keep track of which pins knocked down

2. Probability of spare depends on configuration of the pins remaining

3. Simulate 1000 times

4. Generate a histogram of outcomes

5. Generaate a covariance matrix for first rolls




E[XY]âˆ’(EX)(EY)
```{r}
x <- as_tibble(c(prob_pin_1, prob_pin_2, prob_pin_3, prob_pin_4, prob_pin_5, prob_pin_6, prob_pin_7, prob_pin_8, prob_pin_9, prob_pin_10))
 colnames(x) <- "X"
y <- as_tibble(c(prob_pin_1, prob_pin_2, prob_pin_3, prob_pin_4, prob_pin_5, prob_pin_6, prob_pin_7, prob_pin_8, prob_pin_9, prob_pin_10))
 colnames(y) <- "Y"
data <- full_join(x, y, by = character())

cov(data)
```
[ð¸(ð‘¦âˆ£ð‘§=1)âˆ’ð¸(ð‘¦âˆ£ð‘§=0)]ð‘(1âˆ’ð‘)

```{r}
prob_pin_1


cov_matrix <- matrix(prob_pin_1*(1-prob_pin_1))
```


### Pins Knocked Down

```{r}
frame <- as_tibble(seq(1,10)) %>%
  rename(frame = value)

first_roll <- frame %>%
  mutate(roll_1_pins_down = sample(0:10, 10, replace=TRUE))

roll_1_pins_down_vector <- first_roll$roll_1_pins_down

roll_2_pins_down_vector <- c()

  if(roll_1_pins_down_vector[i] < 5 & 
all(unlist(first_roll$pins_remaining[i])) %in% c("2", "4", "7", "8") | roll_1_pins_down_vector[i] < 5 &  all(unlist(first_roll$pins_remaining[i])) %in% c("3", "6", "9", "10")){
    roll_2_pins_down_vector <- append(roll_2_pins_down_vector, 
                             sample(0:(10 - first_roll$roll_1_pins_down[i]), 1))
  }
  if(roll_1_pins_down_vector[i] < 7 & all(unlist(first_roll$pins_remaining[i])) %in% c("1", "2", "3", "5", "8", "9")){
    roll_2_pins_down_vector <- append(roll_2_pins_down_vector, 
                             sample(0:(10 - first_roll$roll_1_pins_down[i]), 1))
  }


for(i in 1:10){
  if(any(unlist(first_roll$pins_remaining[i])) %in% c("2", "4", "7", "8") &
      any(unlist(first_roll$pins_remaining[i])) %in% c("3", "6", "9", "10") & 
      any(unlist(first_roll$pins_remaining[i])) %in% c("1", "2", "3", "5", "8", "9") == FALSE){
    if(sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")) > sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")))
      {
        roll_2_pins_down_vector <- append(roll_2_pins_down_vector, sample(0:sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")), 1))
    }
    if(sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")) < sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")))
      {
        roll_2_pins_down_vector <- append(roll_2_pins_down_vector, sample(0:sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")), 1))
      }
    if(sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")) == sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")))
      {
        roll_2_pins_down_vector <- append(roll_2_pins_down_vector, sample(0:sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")), 1))
      }
  }
  else{
    roll_2_pins_down_vector <- append(roll_2_pins_down_vector, 
                             sample(0:(10 - first_roll$roll_1_pins_down[i]), 1))
  }
}

first_roll <- first_roll %>%
  mutate(pins_remaining_string = case_when(roll_1_pins_down == 0 ~ pins_remaining_10,
                                    roll_1_pins_down == 1 ~ pins_remaining_9[sample(1:length(pins_remaining_9), 1)],
                                    roll_1_pins_down == 2 ~ pins_remaining_8[sample(1:length(pins_remaining_8), 1)],
                                    roll_1_pins_down == 3 ~ pins_remaining_7[sample(1:length(pins_remaining_7), 1)],
                                    roll_1_pins_down == 4 ~ pins_remaining_6[sample(1:length(pins_remaining_6), 1)],
                                    roll_1_pins_down == 5 ~ pins_remaining_5[sample(1:length(pins_remaining_5), 1)],
                                    roll_1_pins_down == 6 ~ pins_remaining_4[sample(1:length(pins_remaining_4), 1)],
                                    roll_1_pins_down == 7 ~ pins_remaining_3[sample(1:length(pins_remaining_3), 1)],
                                    roll_1_pins_down == 8 ~ pins_remaining_2[sample(1:length(pins_remaining_2), 1)],
                                    roll_1_pins_down == 9 ~ pins_remaining_1[sample(1:length(pins_remaining_1), 1)],
                                    roll_1_pins_down == 10 ~ "")) %>%
  mutate(pins_remaining = strsplit(pins_remaining_string, ','))


  
second_roll <- first_roll %>%
 mutate(roll_2_pins_down = case_when(roll_1_pins_down != 10 
                                     ~ roll_2_pins_down_vector[frame],
                                     roll_1_pins_down == 10 ~ NA_integer_))

third_roll <- second_roll %>%
  mutate(roll_2_pins_down = case_when(frame == 10 & roll_1_pins_down == 10 
                                           ~ sample(0:10, 1),
                                      TRUE ~ roll_2_pins_down),
         roll_3_pins_down = case_when(frame == 10 & roll_1_pins_down == 10 & roll_2_pins_down == 10 
                                           ~ sample(0:10, 1),
                                      frame == 10 & roll_1_pins_down == 10 & roll_2_pins_down != 10 ~ sample(0:(10 - roll_2_pins_down[10]), 1),
                                      frame == 10 & roll_1_pins_down + roll_2_pins_down == 10 ~ sample(0:10, 1),
                                      TRUE ~ NA_integer_))

roll_1_pins_down_vector <- third_roll$roll_1_pins_down
roll_2_pins_down_vector <- third_roll$roll_2_pins_down
roll_3_pins_down_vector <- third_roll$roll_3_pins_down
```


```{r}
firs_roll_numeric <- first_roll$pins_remaining %>%
  mutate(pins_remaining = as.numeric(unlist(first_roll$pins_remaining)))
as.numeric(unlist(first_roll$pins_remaining[i]))

for(i in 1:10){
  count_both <- 0
  count_x <- 0
  count_y <- 0
  temp <- c()
  for(j in 1:10){
    if(!(i %in% as.numeric(unlist(first_roll$pins_remaining[i])))){
      temp <- append(temp, i)
    }
  temp <- as.numeric(unlist(first_roll$pins_remaining[i]))
  if 
  if (x %in% temp){
    count_x <- count_x + sum(temp == x)
  }
  count_both <- count_both + 
}
cov1.1 <- 
```

### Frame Interpretation

```{r}
frame_interpretations <- third_roll %>%
  mutate(frame_interpretation = case_when(roll_1_pins_down == 10 ~ "STRIKE",
                                          roll_1_pins_down + roll_2_pins_down == 10 ~ "SPARE",
                                          TRUE ~ "OPEN"))
```


### Frame Points

```{r}
points <- frame_interpretations %>%
  mutate(frame_points = case_when(frame_interpretation == "OPEN" ~ as.numeric(roll_1_pins_down_vector[frame] + roll_2_pins_down_vector[frame]), # open
                      
                                  frame != 10 & frame_interpretation == "SPARE" ~ 10 + lead(roll_1_pins_down, order_by = frame, n = 1), # spare before 10th frame
                                  
                                  frame == 10 & frame_interpretation == "SPARE" ~ 10 + roll_3_pins_down_vector[frame], # spare on 10th frame
                                  
                                  frame < 9 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) == "STRIKE" ~ 20 + lead(roll_1_pins_down, order_by = frame, n = 2), # 2 strikes in a row before 9th
                                  
                                  frame != 10 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) != "STRIKE" ~ 10 + lead(roll_1_pins_down, order_by = frame, n = 1) + lead(roll_2_pins_down, order_by = frame, n = 1), # strike before 10th frame
                                  
                                  frame == 9 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) == "STRIKE" ~ 20 + lead(roll_2_pins_down, order_by = frame, n = 1), # 2 strikes in a row 9th
                                  
                                  frame = 10 & frame_interpretation == "STRIKE" ~ 10 + roll_2_pins_down_vector[10] + roll_3_pins_down_vector[10], # 10th frame strike
                                  ))
```


### Frame Summaries and Total Score

```{r}
frame_summaries <- points %>%
  select(-pins_remaining) %>%
  kable(col.names = c("Frame", "Pins Remaining", "Roll 1 Pins Down", "Roll 2 Pins Down", "Roll 3 Pins Down", "Frame Interpretation", "Frame Points"), align = "r") %>%
  kableExtra::kable_styling("striped")


total_score <- points %>%
  summarize(total_points = sum(frame_points)) %>%
  kable(col.names = "Total Points", align = "r") %>%
  kableExtra::kable_styling("striped")
```


### Function for Single Iteration of Score Simulation


```{r message = FALSE, warning = FALSE}
sim_1000 <- function() {
  frame <- as_tibble(seq(1,10)) %>%
  rename(frame = value)

first_roll <- frame %>%
  mutate(roll_1_pins_down = sample(0:10, 10, replace=TRUE))

roll_1_pins_down_vector <- first_roll$roll_1_pins_down

roll_2_pins_down_vector <- c()

first_roll <- first_roll %>%
  mutate(pins_remaining_string = case_when(roll_1_pins_down == 0 ~ pins_remaining_10,
                                    roll_1_pins_down == 1 ~ pins_remaining_9[sample(1:length(pins_remaining_9), 1)],
                                    roll_1_pins_down == 2 ~ pins_remaining_8[sample(1:length(pins_remaining_8), 1)],
                                    roll_1_pins_down == 3 ~ pins_remaining_7[sample(1:length(pins_remaining_7), 1)],
                                    roll_1_pins_down == 4 ~ pins_remaining_6[sample(1:length(pins_remaining_6), 1)],
                                    roll_1_pins_down == 5 ~ pins_remaining_5[sample(1:length(pins_remaining_5), 1)],
                                    roll_1_pins_down == 6 ~ pins_remaining_4[sample(1:length(pins_remaining_4), 1)],
                                    roll_1_pins_down == 7 ~ pins_remaining_3[sample(1:length(pins_remaining_3), 1)],
                                    roll_1_pins_down == 8 ~ pins_remaining_2[sample(1:length(pins_remaining_2), 1)],
                                    roll_1_pins_down == 9 ~ pins_remaining_1[sample(1:length(pins_remaining_1), 1)],
                                    roll_1_pins_down == 10 ~ "")) %>%
  mutate(pins_remaining = strsplit(pins_remaining_string, ','))


  



for(i in 1:10){
  if(any(unlist(first_roll$pins_remaining[i])) %in% c("2", "4", "7", "8") &
      any(unlist(first_roll$pins_remaining[i])) %in% c("3", "6", "9", "10") & 
      any(unlist(first_roll$pins_remaining[i])) %in% c("1", "2", "3", "5", "8", "9") == FALSE){
    if(sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")) > sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")))
      {
        roll_2_pins_down_vector <- append(roll_2_pins_down_vector, sample(0:sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")), 1))
    }
    if(sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")) < sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")))
      {
        roll_2_pins_down_vector <- append(roll_2_pins_down_vector, sample(0:sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")), 1))
      }
    if(sum(unlist(first_roll$pins_remaining[i]) %in% c("3", "6", "9", "10")) == sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")))
      {
        roll_2_pins_down_vector <- append(roll_2_pins_down_vector, sample(0:sum(unlist(first_roll$pins_remaining[i]) %in% c("2", "4", "7", "8")), 1))
      }
  }
  else{
    roll_2_pins_down_vector <- append(roll_2_pins_down_vector, 
                             sample(0:(10 - first_roll$roll_1_pins_down[i]), 1))
  }
}

second_roll <- first_roll %>%
 mutate(roll_2_pins_down = case_when(roll_1_pins_down != 10 
                                     ~ roll_2_pins_down_vector[frame],
                                     roll_1_pins_down == 10 ~ NA_integer_))

third_roll <- second_roll %>%
  mutate(roll_2_pins_down = case_when(frame == 10 & roll_1_pins_down == 10 
                                           ~ sample(0:10, 1),
                                      TRUE ~ roll_2_pins_down),
         roll_3_pins_down = case_when(frame == 10 & roll_1_pins_down == 10 & roll_2_pins_down == 10 
                                           ~ sample(0:10, 1),
                                      frame == 10 & roll_1_pins_down == 10 & roll_2_pins_down != 10 ~ sample(0:(10 - roll_2_pins_down[10]), 1),
                                      frame == 10 & roll_1_pins_down + roll_2_pins_down == 10 ~ sample(0:10, 1),
                                      TRUE ~ NA_integer_))



roll_1_pins_down_vector <- third_roll$roll_1_pins_down
roll_2_pins_down_vector <- third_roll$roll_2_pins_down
roll_3_pins_down_vector <- third_roll$roll_3_pins_down

frame_interpretations <- third_roll %>%
  mutate(frame_interpretation = case_when(roll_1_pins_down == 10 ~ "STRIKE",
                                          roll_1_pins_down + roll_2_pins_down == 10 ~ "SPARE",
                                          TRUE ~ "OPEN"))

points <- frame_interpretations %>%
  mutate(frame_points = case_when(frame_interpretation == "OPEN" ~ as.numeric(roll_1_pins_down_vector[frame] + roll_2_pins_down_vector[frame]), # open
                      
                                  frame != 10 & frame_interpretation == "SPARE" ~ 10 + lead(roll_1_pins_down, order_by = frame, n = 1), # spare before 10th frame
                                  
                                  frame == 10 & frame_interpretation == "SPARE" ~ 10 + roll_3_pins_down_vector[frame], # spare on 10th frame
                                  
                                  frame < 9 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) == "STRIKE" ~ 20 + lead(roll_1_pins_down, order_by = frame, n = 2), # 2 strikes in a row before 9th
                                  
                                  frame != 10 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) != "STRIKE" ~ 10 + lead(roll_1_pins_down, order_by = frame, n = 1) + lead(roll_2_pins_down, order_by = frame, n = 1), # strike before 10th frame
                                  
                                  frame == 9 & frame_interpretation == "STRIKE" & lead(frame_interpretation, order_by = frame, n = 1) == "STRIKE" ~ 20 + lead(roll_2_pins_down, order_by = frame, n = 1), # 2 strikes in a row 9th
                                  
                                  frame = 10 & frame_interpretation == "STRIKE" ~ 10 + roll_2_pins_down_vector[10] + roll_3_pins_down_vector[10], # 10th frame strike
                                  ))
total_score <- points %>%
  summarize(total_points = sum(frame_points)) %>%
  select(total_points) %>%
  pull()

return(total_score) 

}

```


### Repeat to Get 1000 Scores

```{r message = FALSE, warning = FALSE}
scores <- c()

for(i in 1:1000){
  scores <- append(scores, sim_1000())
}
```

### Histogram of Outcomes

```{r}
scores_tibble <- as_tibble(scores)

ggplot(scores_tibble, aes(x = value)) +
  geom_histogram(bins = 30, fill = "gray", color = "black") +
  labs(title = "Distribution of 1000 Bowling Score Simulations",
       x = "Bowling Game Score",
       y = "Frequency") +
  theme_minimal()
  
```
